Issues with Dynamical Measurements Across the Disk

### 2019-03-18 paper/ms.tex
first revisions to start of introduction

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution in the
era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
major merger
\citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
a phase-space spiral \citep{2018Natur.561..360A} indicating either local
substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
buckling \citep{2018arXiv181109205K}, a gap indicative of a dark matter
subtructure in GD1 \citep{2018arXiv181103631B}, among others.

These discoveries are all indicative that the Milky Way departs from
axisymmetry and is undergoing phase mixing and dynamical interactions across a
range of spatial and temporal scales. Underlying the quantitative description
of the mechanisms at work that give rise to these signatures is the assumption
of a Galactocentric coordinate system \citep{2008gady.book.....B} which can
both precisely and accurately be defined and measured. In angular coordinates,
the center of the galaxy is taken to be the location of
Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
motions near Sgr~A\textsuperscript{*}, the distance to the center is taken to
be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The vertical height is found
by identifying where the stellar positions and velocities reach a maximum
(effectively the median height of all disk stars), and is usually taken to be
$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}. A
pre-{\em Gaia} review of these measurements is given by
\citet{2016ARAA..54..529B}.

Once a Galactocentric coordinate system has been established and a 6D phase
space measurement of a star has been made, it is desirable to convert this
measurement into action space. Actions are quantities which describe the orbit
of a star and are conserved quantities under some assumptions which we
describe next. They are the typical integral of the Hamiltonian in cylindrical
coordinates and are given by,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
\eeq
where $i=R,z,\phi$.\footnote{We typically write $J_r$ instead of $J_R$ for
aesthetic reasons.} Note that $J_{\phi} \equiv L_z$. In a slowly-evolving
axisymmetric potential, the actions of stars' orbits are conserved quantities
\citep{2008gady.book.....B,2014RvMP...86....1S}. The non-axisymmetries of a
more realistic potential induce time evolution in the {\em measured} actions
--- the true orbital invariants (if they exist) are invariant by definition.

@@

\begin{deluxetable*}{cccccc}
\tablecaption{Stellar disk scale heights of the real Milky Way and the
simulated galaxies considered in this work.\label{tab:scale_height}}
\tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} & \colhead{stellar thin disk} & \colhead{stellar thick disk} & \colhead{cold gas softening length} & \colhead{stellar softening length} \\ 
\colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} } 
\startdata
MW\tablenotemark{b} & 40 & 300 & 900 & 0 & 0 \\
\mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\
\mf\tablenotemark{c} & 360 & 440 & 1280 & 12.5 & 11.2 \\
\mm\tablenotemark{c} & 250 & 290 & 1030 &10.3 & 11.2 \\
\enddata
% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}
\tablenotetext{a}{$T<1000$ K}
\tablenotetext{b}{\citet{2008ApJ...673..864J,2016ARAA..54..529B}}
\tablenotetext{c}{\citet{2018arXiv180610564S}}
\tablenotetext{d}{The azimuthally averaged gas vertical density profile in
\mi{} is nearly constant to this height, though individual regions show smaller
scale heights and dense clouds.}
\end{deluxetable*}

@@

Figure~\ref{fig:Rn_mc} shows that azimuthal midplane fluctuations will limit
the volume over which actions can be used to match stars to open clusters on
thin-disk-like orbits (\thincolor), depending on their mass. For
thick-disk-like orbits (\thickcolor), the tolerances are much more forgiving:
stars' actions can be compared over the entire Solar circle for nearly all
cluster masses, except at the very low-mass end in \mi . The neighborhood
around $\sim100\Msun$ clusters can be as small as $\sim300\,\pc$ but up to
$\sim1.2\,\kpc$ for \mi{} and \mf{}, though the situation is marginally better
for \mm{} with values ranging from $\sim0.8\,\kpc$ to $\sim3\,\kpc$. The
neighborhood values steadily increase with $m_c$, topping out at $\sim5\,\kpc$
for $\sim10^4\,\Msun$ clusters in
\mi{} and \mf{}. In \mm{}, the neighborhood size reaches the maximum allowed of
$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. This figure also
underlines that the degree to which such an analysis will succeed depends
crucially on the expected degree of midplane fluctuations in the real Galaxy.
If the Milky Way is more like \mi , the range of action-space analysis
techniques like this is far more limited than if it more closely resembles \mm
.

>>>>

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution in the
era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
major merger
\citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation by a dark matter
subtructure in the tidal stream GD1 \citep{2018arXiv181103631B}.
These discoveries all indicate that the Milky Way's stellar distribution, which demonstrably departs significantly from axisymmetry, is undergoing phase mixing and dynamical interactions across a range of spatial and temporal scales.

Underlying the quantitative analysis
of many of the mechanisms at work that give rise to these signatures is the assumption of a global, axisymmetric Galactocentric coordinate system \citep{2008gady.book.....B} in which the Sun's position (and therefore the origin of the coordinate system) can be defined and measured both precisely and accurately. In this system,
the center of the galaxy is usually taken to be the location of the central supermassive black hole,
Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
motions near Sgr~A\textsuperscript{*}, the distance from the Sun to the Galactic center is taken to
be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The Sun's vertical distance from the Galactic midplane, the other independent measurement that sets this coordinate system, can be determined by identifying where the stellar density and velocities reach a maximum
(effectively the median height of all disk stars), and is usually taken to be
$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}. Another strategy is to use the cold gas or HII regions in the disk to define the Galactic midplane, leading to slightly values ($\sim 5\,\pc$) for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A
pre-{\em Gaia} review of these measurements is given by
\citet{2016ARAA..54..529B}.

Once a Galactocentric coordinate system has been established and a 6D phase
space measurement of a star has been made, it is often desirable to convert this
measurement into action space to concisely summarize its projected orbit, model the stellar distribution function, or find stars with similar dynamical properties. Actions are conserved quantities which describe the orbit
of a star under the assumption of regular, bound orbits in a system where the equations of motion are separable in a particular coordinate system. They are the typical integral of the Hamiltonian in cylindrical
coordinates and are given by,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
\eeq
where $i=R,z,\phi$.\footnote{We typically write $J_r$ instead of $J_R$ for
aesthetic reasons.} Note that $J_{\phi} \equiv L_z$. In a slowly-evolving
axisymmetric potential, the actions of stars' orbits are conserved quantities
\citep{2008gady.book.....B,2014RvMP...86....1S}. The non-axisymmetries of a
more realistic potential induce time evolution in the {\em measured} actions
--- the true orbital invariants (if they exist) are invariant by definition.

@@

\begin{deluxetable*}{cccccc}
\tablecaption{Stellar disk scale heights of the real Milky Way and the
simulated galaxies considered in this work.\label{tab:scale_height}}
\tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} & \colhead{stellar thin disk} & \colhead{stellar thick disk} & \colhead{cold gas softening length} & \colhead{stellar softening length} \\ 
\colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} } 
\startdata
MW\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
\mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\
\mf\tablenotemark{c} & 360 & 440 & 1280 & 12.5 & 11.2 \\
\mm\tablenotemark{c} & 250 & 290 & 1030 &10.3 & 11.2 \\
\enddata
% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}
\tablenotetext{a}{$T<1000$ K}
\tablenotetext{b}{\citet{2008ApJ...673..864J,2016ARAA..54..529B}}
\tablenotetext{c}{\citet{2018arXiv180610564S}}
\tablenotetext{d}{The azimuthally averaged gas vertical density profile in
\mi{} is nearly constant to this height, though individual regions show smaller
scale heights and dense clouds.}
\end{deluxetable*}

@@

Figure~\ref{fig:Rn_mc} shows that azimuthal midplane fluctuations will limit
the volume over which actions can be used to match stars to open clusters on
thin-disk-like orbits (\thincolor), depending on their mass. For
thick-disk-like orbits (\thickcolor), the tolerances are much more forgiving:
stars' actions can be compared over the entire Solar circle for nearly all
cluster masses, except at the very low-mass end in \mi . The neighborhood
around $\sim100\Msun$ clusters can be as small as $\sim300\,\pc$ but up to
$\sim1.2\,\kpc$ for \mi{} and \mf{}, though the situation is marginally better
for \mm{} with values ranging from $\sim0.8\,\kpc$ to $\sim3\,\kpc$. The
neighborhood values steadily increase with $m_c$, topping out at $\sim5\,\kpc$
for $\sim10^4\,\Msun$ clusters in
\mi{} and \mf{}. In \mm{}, the neighborhood size reaches the maximum allowed of
$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. This figure also
underlines that the degree to which such an analysis will succeed depends
crucially on the expected degree of midplane fluctuations in the real Galaxy.
If the Milky Way is more like \mi , the range of action-space analysis
techniques like this is far more limited than if it more closely resembles
\mm{}.
